---
title: "Prediction of Type of Exercise Based on Accelerometer Data"
output: html_document
---

### Introduction

This is a brief summary of the model built to predict the type of exercise based on activity.

### Loading Data and Pre-Processing

```{r, cache=TRUE}
pml_test <- read.csv("pml-testing.csv")
pml_train <- read.csv("pml-training.csv")
```


Dimensions of each dataset:
```{r}
dim(pml_test)
dim(pml_train)
```


Check for "NA" values in the datasets. As shown below, almost every row has an "NA" value.
```{r}
sum(complete.cases(pml_test))
sum(complete.cases(pml_train))
```


Remove columns that mostly have "NA" and missing values.
```{r, cache=TRUE}
cols <- colSums(is.na(pml_test)) != nrow(pml_test)
pml_test <- pml_test[, cols]
pml_train <- pml_train[, cols]
```


Check to confirm NA values no longer appear in the datasets.
```{r}
sum(complete.cases(pml_test)) == nrow(pml_test)
sum(complete.cases(pml_train)) == nrow(pml_train)
```


The first 7 columns are omitted as they mostly identify the user or they were not data acquired by the accelerometer.
```{r}
names(pml_test)
```


We remove these columns.
```{r, cache=TRUE}
pml_train <- pml_train[,-c(1:7)]
pml_test <- pml_test[,-c(1:7)]
```


### Model

We use Random Forest to build the prediction model.
```{r, cache=TRUE}
library(randomForest)
set.seed(12345)
modFit <- randomForest(classe ~ ., data=pml_train)
```

Since we used Random Forest, we do not need a separate test/cross validation set to get an unbiased test set error.[1] We can see that the model does a fairly good job of predicting the outcome with an out of sample error rate of `r round(modFit$err.rate[500,1]*100,2)`%.

```{r, cache=TRUE}
modFit
```

The confusion matrix shows that very few were cases were misclassified.

This model accurately predicts the activities for the 20 cases in the test set. This can be verified by the autograder on Coursera.
```{r, cache=TRUE}
predict(modFit, pml_test)
```

### Links
[1] http://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm#ooberr